%! Author = evert
%! Date = 21-7-2024

\section{Research goals}
The main research goal behind this proposal is, finding answers to:

\textbf{How does the profit for delivery companies vary when using independent contractors versus hiring deliverers as employees with an hourly wage? }

An instrument will be built in the form of an Agent Based Model.
Calling it an instrument does not mean it is not an important deliverable, as most time and energy will be spent in creating it.

Thus, the second goal of the research is: building an Agent Based Model (ABM) that reflect a world where 3 types of agents exist: restaurants, customers and deliverers.

The initial basic requirements of this model are:

\begin{itemize}
    \item   The world consists op a grid of squares, some squares will be blocks of buildings and others represent the streets.
    \item   On this grid some predetermined restaurants and customers exist.
    \item   Customers will order food at a restaurant, the restaurant will prepare the food en will ask for a deliverer.
    \item   A deliverer will claim the delivery, move to the restaurant, pick the food up and move to the customer.
    \item   Agents may leave the world, i.e.,choose another delivery company, when unhappy.
    \item   Unhappiness will occur:
    \begin{itemize}
        \item   for customers when the food is not or too late delivered,
        \item   for restaurants when the food is not picked up or no orders are being placed,
        \item   for deliverers when they don't make any money.
    \end{itemize}
    \item    Calculating the profit for the delivery company.
    \item    The simulation consists of discreet steps, in which all agents simultaneous can do one thing.
    Some examples are:
    \begin{itemize}
        \item  move to the next square
        \item  place an order at a restaurant
        \item  do nothing
        \item  decide to take an order
    \end{itemize}
    \item   Behavior of the agents is rule based, these have to be programmed.
\end{itemize}

More requirements will be found during development, and some will be dropped.

To summarize the design problem: to improve insight into profitability in whether to employ deliverers,
an ABM will be created in NetLogo that replicates a real world food delivery situation and satisfies the above-listed requirements,
so that stakeholders, including the author, can use this to make an unbiased analysis.
